Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-10 inc/design_considerations/considerations.tex
bugfix
more work on single_photon photoemission
killing more TODOs

In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
However, microscopy is an experimental science, and designing a UEM is no less hands-on.
%TODO avoid repitition? "When doing so, special ..."?
When designing a UEM system, special care must be given to the creation and propagation of the electron beam.
%TODO define single-shot and transverse emittance here or rephrase?
In the case of single-shot UEM, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, the transverse emittance of the beam must be minimized.

@@

\section{Beam Quality and Resolution Limit}

A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
\begin{equation} \label{eq:emittance_definition}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > - \left < x p_x \right >^2 } \,\text{,}
\end{equation}
where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_x$ are the transverse position and momentum respectively.
For the usual case of axially symmetric electron pulses, we may use the simplified form
\begin{equation} \label{eq:emittance_definition}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{.}
\end{equation}
By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throughout pulse propagation (for free-space and perfect electron-optical elements), and therefore is related directly to the initial photoemission properties of the system.
For a laser pulse of the form of \ref{eq:laser_form} we have simply
\begin{equation} \label{eq:emittance_definition_w}
  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_x^2 \right >}{2} } \,\text{.}
\end{equation}
Note that imperfect electron-optical elements may cause ``emittance growth'' \cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.

>>>>

In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
However, since microscopy is an experimental science, we are also designing and building a prototype UEM system at UIC.
In doing so we have decided that special care must be given to the creation and propagation of the electron beam.
In the case of single-shot UEM, in which the image of the sample is acquired using only one electron pulse, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, that the ``normalized transverse emittance,'' a beam quality figure of merit which will be defined later, must be minimized.

@@

\section{Beam Quality and Resolution Limit} \label{sec:must_reduce_transverse_momentum}

A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
\begin{equation}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > - \left < x p_x \right >^2 } \,\text{,}
\end{equation}
where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_x$ are the transverse position and momentum respectively.
For the usual case of axially symmetric electron pulses, we may use the simplified form
\begin{equation} \label{eq:emittance_definition}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{.}
\end{equation}
By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throughout pulse propagation (for free-space and perfect electron-optical elements), and therefore is related directly to the initial photoemission properties of the system.
For a laser pulse of the form of \ref{eq:laser_form} we have simply
\begin{equation} \label{eq:emittance_definition_w}
  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_x^2 \right >}{2} } \,\text{.}
\end{equation}
Note that imperfect electron-optical elements may cause ``emittance growth'' \cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.

### 2013-04-10 inc/extension/external_forces/formal.tex
killing more TODOs

The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}.
As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.%, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
%TODO consider adding a replacement sentance like: "It would be interesting to employ ..." or "A future goal is to ..."

>>>>

The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}.
As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.

### 2013-04-10 inc/extension/external_forces/practical.tex
killing more TODOs

The mathematical formalism developed in Section \ref{sec:external_forces} is only an abstract mechanism for including external forces.
To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression and acceleration %TODO refs
) and DC acceleration regions (i.e. electron gun dynamics).

@@

%TODO perhaps this shouldn't be a redefinition, but some other function
Of course, physical electron-optical elements occupy only a finite space in the column, and thus their influence on the pulse should be limited to a region near the element itself.
To localize the mathematical result shown in Section \ref{sec:external_forces}, it is convenient to redefine a particular $M_{\alpha}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
\begin{equation} \label{eq:reg_of_influence}
  M_{\alpha} \to M_{\alpha} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
\end{equation}
where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the axial length of the lens and $n$ is the super-Gaussian order parameter.
The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $; the kinetic equations are simply defined for the position of the peak of the pulse.

@@

A simple magnetic lens is just a solenoid whose axis is collinear with the axis of the microscope column.
%Initially one would think that electrons moving parallel to this axis would not be affected by the magnetic field lines in the solenoid, however the contribution of the fringing field must be taken into consideration.
As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
This ``tumbling'' around the axis, along with the axial magnetic field, is what causes the inward-directed force.

The full description of the motion of an electron (in cylindrical coordinates $r, \theta, z$) in an azimuthally symmetric magnetic field ($\vec{B} = B_r \hat{r} + B_z \hat{z}$) is easily derived \cite{el-kareh_electron_1970},
\begin{subequations} \label{eq:lens_eq_of_motion}
\begin{gather}
  m \ddot{r} = -q B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
  \frac{d}{dt} ( m r^2 \dot{\theta} ) = q r \dot{r} B_z - q r B_r \dot{z} \\
  m \ddot{z} = q B_r r \dot{\theta}
\end{gather}
\end{subequations}
where each dot denotes a time derivative.
Attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
Compounding the problem, the full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$, again in cylindrical coordinates, 
\begin{subequations} \label{eq:field_of_loop}
\begin{gather}
  \frac{B_z}{\mu} = H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
  \frac{B_r}{\mu} = H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
  Q \equiv (a+r)^2 + z^2 \\
  k \equiv \sqrt{ 4 a r / Q }
\end{gather}
\end{subequations}
%%TODO I can't decide whether the Q and k definitions should be numbered or not
%where
%\begin{gather*}
%  Q \equiv (a+r)^2 + z^2 \\
%  k \equiv \sqrt{ 4 a r / Q }
%\end{gather*}
where $\mu$ is the local magnetic permability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
In the vacuum, $\mu$ is simply the usual free space permeability $\mu_0$; in Section \ref{sec:mag_lens}, we will see how this permeability can be used to improve the performance of magnetic lenses.
An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.

@@

For localizing the region of influence of the lens, I have generally modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, one adds a term like
\begin{equation}
  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
\end{equation}
to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.
%TODO include discussion and plot of beam size versus aberrations

@@

The AG model very accurately mimics the behavior of the kinematic model for an appropriate choice of the free parameter $M_{\smallT}^{mag}$.
In fact, when the rays are weighted by a Gaussian of width $w$ the minimum ``radius'' of the beam is 100$\mu$m, while the AG model has a minimum HW1/eM width of 73$\mu$m.
(The kinematic model actually overestimates the width as there are two rays at the outer convergence zone for every one which converges at the axis.) %TODO footnote?

@@

The two main effects contributing to the transverse lens of a $\text{TM}_{010}$-mode RF cavity, its magnetic field component and electric fringe field aperture effects, may also be included within the linear force approximation.
The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $ \cite{kim_rf_1989}, where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs \cite{mcdonald_design_1988}). 
For the case of an evacuated cavity, these two contributions result in a net force,
\begin{equation} \label{eq:RF_Force_r}
  \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
    \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
   \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
\end{equation}
In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
%TODO examples shown below?
For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

To summarize, the additional terms are which are added to model the internal pulse dynamics arising from a single TM$_{010}$-mode cylindrical RF cavity are
\begin{gather}
  M_{z}^{RF} = \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \\
  \begin{aligned}
  M_{\smallT}^{RF} = & e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \right. \\ \times & \left. \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ]
  \end{aligned}
\end{gather}
Additionally, since an RF cavity can act as a pulse accelerator (as seen in the leading term of \ref{eq:RF_Efield_approx}) an additional force term
\begin{equation}
  F_{z}^{RF} = e E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
\end{equation}
should be added to the equation of motion governing the propagation of the pulse down the column.
% excluded example section from JAP paper as it will need new graphics

>>>>

The mathematical formalism developed in Section \ref{sec:external_forces} is only an abstract mechanism for including external forces.
To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression \cite{oudheusden_electron_2007} and acceleration %TODO refs
) and DC acceleration regions (i.e. electron gun dynamics).

@@

Of course, physical electron-optical elements occupy only a finite space in the column, and thus their influence on the pulse should be limited to a region near the element itself.
To localize the mathematical result shown in Section \ref{sec:external_forces}, it is convenient to redefine a particular $M_{\alpha}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
\begin{equation} \label{eq:reg_of_influence}
  M_{\alpha} \to M_{\alpha} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
\end{equation}
where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the axial length of the lens and $n$ is the super-Gaussian order parameter.
The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $; the kinetic equations are simply defined for the position of the peak of the pulse.

@@

A simple magnetic lens is just a solenoid whose axis is collinear with the axis of the microscope column.
As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
This ``tumbling'' around the axis, along with the axial magnetic field, is what causes the inward-directed force.

The full description of the motion of an electron (in cylindrical coordinates $r, \theta, z$) in an azimuthally symmetric magnetic field ($\vec{B} = B_r \hat{r} + B_z \hat{z}$) is easily derived \cite{el-kareh_electron_1970},
\begin{subequations} \label{eq:lens_eq_of_motion}
\begin{gather}
  m \ddot{r} = -q B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
  \frac{d}{dt} ( m r^2 \dot{\theta} ) = q r \dot{r} B_z - q r B_r \dot{z} \\
  m \ddot{z} = q B_r r \dot{\theta}
\end{gather}
\end{subequations}
where each dot denotes a time derivative.
Attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
Compounding the problem, the full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$, again in cylindrical coordinates, 
\begin{subequations} \label{eq:field_of_loop}
\begin{gather}
  \frac{B_z}{\mu} = H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
  \frac{B_r}{\mu} = H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
  Q \equiv (a+r)^2 + z^2 \\
  k \equiv \sqrt{ 4 a r / Q }
\end{gather}
\end{subequations}
where $\mu$ is the local magnetic permability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
In the vacuum, $\mu$ is simply the usual free space permeability $\mu_0$; in Section \ref{sec:mag_lens}, we will see how this permeability can be used to improve the performance of magnetic lenses.
An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.

@@

For localizing the region of influence of the lens, I have generally modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, one adds a term like
\begin{equation}
  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
\end{equation}
to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.

@@

The AG model very accurately mimics the behavior of the kinematic model for an appropriate choice of the free parameter $M_{\smallT}^{mag}$.
In fact, when the rays are weighted by a Gaussian of width $w$ the minimum ``radius'' of the beam is 100$\mu$m, while the AG model has a minimum HW1/eM width of 73$\mu$m.
(The kinematic model actually overestimates the width as there are two rays at the outer convergence zone for every one which converges at the axis.)

@@

The two main effects contributing to the transverse lens of a $\text{TM}_{010}$-mode RF cavity, its magnetic field component and electric fringe field aperture effects, may also be included within the linear force approximation.
The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $ \cite{kim_rf_1989}, where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs \cite{mcdonald_design_1988}). 
For the case of an evacuated cavity, these two contributions result in a net force,
\begin{equation} \label{eq:RF_Force_r}
  \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
    \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
   \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
\end{equation}
In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

To summarize, the additional terms are which are added to model the internal pulse dynamics arising from a single TM$_{010}$-mode cylindrical RF cavity are
\begin{gather}
  M_{z}^{RF} = \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \\
  \begin{aligned}
  M_{\smallT}^{RF} = & e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \right. \\ \times & \left. \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ]
  \end{aligned}
\end{gather}
Additionally, since an RF cavity can act as a pulse accelerator (as seen in the leading term of \ref{eq:RF_Efield_approx}) an additional force term
\begin{equation}
  F_{z}^{RF} = e E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
\end{equation}
should be added to the equation of motion governing the propagation of the pulse down the column.

### 2013-04-10 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
killing more TODOs

Not only is this an interesting result, but it can be both instructive and computationally useful.
By evaluating $\Gamma_{\alpha}$ at some time (presumably $t=0$) and rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, we can rewrite \ref{eq:ag_original} as 
\begin{subequations} \label{eq:ag_big_gamma}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) \, \text{.} \label{eq:ag_big_gamma_gamma}
\end{gather}
\end{subequations}
These four equations now contain all of the information previously contained in six (\ref{eq:ag_original}).
Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
This will become more apparent when, in future sections,
%TODO include internal reference
it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.} \cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
%TODO include more commentary on $\varepsilon_{\smallT}$?
Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.
This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems \cite{oshea_reversible_1998}.
As will be seen in Section \ref{extension_limitations}, the AG model will, of necessity, only consider perfect lenses.

>>>>

Not only is this an interesting result, but it can be both instructive and computationally useful.
By evaluating $\Gamma_{\alpha}$ at some time (presumably $t=0$) and rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, we can rewrite \ref{eq:ag_original} as 
\begin{subequations} \label{eq:ag_big_gamma}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) \, \text{.} \label{eq:ag_big_gamma_gamma}
\end{gather}
\end{subequations}
These four equations now contain all of the information previously contained in six (\ref{eq:ag_original}).
Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
%This will become more apparent when, in future sections, it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.
This becomes apparent when noting that if a momentum chirp $\gamma_{\alpha}$ should ever reach zero (e.g. as it does at the focal point of a lens) there is still a non-Coulombic term driving the $\frac{d\gamma_{\alpha}}{dt}$, as there surely must be when the local momentum variance $\eta_{\alpha}$ is not zero.

We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.} \cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.
This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems \cite{oshea_reversible_1998}.
As will be seen in Section \ref{extension_limitations}, the AG model will, of necessity, only consider perfect lenses.
